---
title: "George Mamvura Submission 2"
author: "George Mamvura"
date: "29/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(readr)
library(openxlsx)
library(dplyr)
library(janitor)



data <- readWorkbook("Insurance Premium Default-Dataset.xlsx", sheet = 1)



data <- clean_names(data)

data_plot <- data%>%mutate_at(vars("id","marital_status","accomodation", "sourcing_channel", "residence_area_type","default"), as.factor)%>%mutate(Months = floor(age_in_days/30))







```

######create new variables for the data####

steps:- 

1. Create variable for late payment 
2. Create age in years varible
4. Create premium bins 
5. Create has dependents 
6. Create owns vehicle variable
7. Create income bins so you know what income bin the people are in 



```{r}


new_data <- data%>%mutate(age_in_years = floor(age_in_days/365),
              late_payment = if_else((count_3_6_months_late == 0 & count_6_12_months_late== 0 &  count_more_than_12_months_late == 0 ), 0,1),
              
              )%>%
       mutate(late_payment = as.factor(late_payment),
              marital_status = as.factor(marital_status), 
              accomodation = as.factor(accomodation),
              sourcing_channel = as.factor(sourcing_channel),
              residence_area_type = as.factor(residence_area_type),
              id = as.factor(id),
              default = as.factor(default))%>%
              rename(no_default = default)%>%
              select(-age_in_days)


```
- everyone has a vehicle and dependents



```{r}


library(ggplot2)
ggplot(new_data, aes(x=no_default, y=premium, fill=no_default)) +
  geom_violin(position=position_dodge(1))
#########look at the income###
ggplot(new_data, aes(x=no_default, y=income, fill=no_default)) +
  geom_violin(position=position_dodge(1))

```
######create income bands in the data#############

Looking at the income values and the premium values and how they vary
```{r}


new_data %>%
  arrange(desc(income)) %>%
  mutate(income_corrected = signif(income, digits = 1),
         premium_corrected = signif(premium, digits = 1))%>%
  group_by(income_corrected,premium_corrected,no_default)%>%
  tally()%>%
  ungroup()%>%
  mutate(index = row_number())%>%
  #mutate(country = factor(country, country)) %>%
  ggplot(aes(x= index , y= income_corrected, size=n, color=no_default)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 29), name="Population (M)")



premiums <- new_data %>%
  #arrange(desc(income)) %>%
  mutate(index = row_number())%>%
  #mutate(country = factor(country, country)) %>%
  ggplot(aes(x= index , y= income, color=no_default)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 29), name="Population (M)")+
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()


premiums_paid <- new_data %>%
  #arrange(desc(income)) %>%
  mutate(index = row_number())%>%
  #mutate(country = factor(country, country)) %>%
  ggplot(aes(x= index , y= premium, color=no_default)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 29), name="Population (M)")+
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()


new_data %>%
  #arrange(desc(income)) %>%
  mutate(index = row_number())%>%
  #mutate(country = factor(country, country)) %>%
  ggplot(aes(x= risk_score , y= perc_premium_paid_by_cash_credit, color=no_default)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 29), name="Population (M)")+
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()


new_data$premium


new_data %>%
  #arrange(desc(income)) %>%
  mutate(index = row_number())%>%
  mutate(income_corrected = signif(income, digits = 1),
         premium_corrected = signif(premium, digits = 1),
         perc_premium_paid_by_cash_credit = signif(perc_premium_paid_by_cash_credit, digits = 1)
         )%>%
  group_by(perc_premium_paid_by_cash_credit,no_default)%>%
  tally()%>%
  ungroup()%>%
  mutate(index = row_number())%>%
  ggplot(aes(x= index , y= perc_premium_paid_by_cash_credit,size = n,  color=no_default)) +
    geom_point(alpha=0.5) +
    scale_size(range = c(.1, 29), name="Population (M)")+
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()




```

```{r}

DataExplorer::create_report(new_data,
                            #output_format = "html",
                            output_file = "Insurance_report_explore.html",
                            output_dir = getwd(), 
                            report_title = "Data `exploration for Insurance Dataset",
                            y = "default"
                            
                            )



```

###check the values they have 

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ggthemes)


pdf(file = "bar_plots_2.pdf", paper = "special",width=9,height=7)
new_data%>%ungroup()%>%
               select( no_default,  late_payment, age_in_years)%>%
  gather(x, y, late_payment:age_in_years) %>%
  count(no_default, x, y) %>%
  ggplot(aes(x = y, y = n, fill = no_default, color = no_default)) +
    facet_wrap(~ x, ncol = 3, scales = "free") +
    geom_bar(stat = "identity", alpha = 1) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "top") +
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()
dev.off()

```


```{r}


pdf(file = "density_plots.pdf", paper = "special",width=9,height=7)
new_data%>%select(no_default, colnames(new_data%>%select_if(is.double))) %>%
            select(no_default, perc_premium_paid_by_cash_credit, age_in_years, income, risk_score, premium)%>%
  gather(x, y, perc_premium_paid_by_cash_credit:premium) %>%
  ggplot(aes(x = y, fill = no_default, color = no_default)) +
    facet_wrap(~ x, ncol = 3, scales = "free") +
    geom_density(alpha = 1, adjust = 3) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "top") +
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()
dev.off()


```



```{r}
DataExplorer::create_report(data_plot,
                            #output_format = "html",
                            output_file = "Insurance_report.html",
                            output_dir = getwd(), 
                            report_title = "Data `exploration for Insurance Dataset",
                            y = "default"
                            
                            )

```


#######SET THE VALUES###




```{r}
#install.packages("Boruta")
library(Boruta)

set.seed(123)
boruta.train <- Boruta(no_default~.-id, data = new_data, doTrace = 2)
print(boruta.train)

`%ni%` <- Negate(`%in%`)

attStats(boruta.train)%>%
  mutate(names = rownames(attStats(boruta.train)))%>%
  filter(names %ni% c("accomodation", "marital_status", "no_of_dep","veh_owned", "residence_area_type"  ) )%>%
  mutate(new_names = factor(names,levels = c("risk_score", "no_of_premiums_paid", "perc_premium_paid_by_cash_credit", "income",
                                             "age_in_years", "count_6_12_months_late", "premium", "late_payment", "count_3_6_months_late",
                                             "sourcing_channel", "count_more_than_12_months_late" )))%>%
  arrange(desc(meanImp))%>%
  ggplot(aes(x = new_names, y = meanImp), ) +
    geom_bar(stat = "identity", alpha = 1, color='skyblue',fill='steelblue') +
    scale_color_tableau() +
    scale_fill_tableau()+
    theme_minimal()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          legend.position = "top")
  

```


####variable selection#########
```{r}

pdf("boruta_train.pdf", width = 14, height = 8, paper = "special")

plot(boruta.train, xlab = "", xaxt = "n")

lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)

boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])

names(lz) <- colnames(boruta.train$ImpHistory)

Labels <- sort(sapply(lz,median))

axis(side = 1,las=2,labels = names(Labels),

at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.7)

dev.off()

```


```{r}

library(lattice)

library(rpart)

#library(brglm)

#install.packages("brglm")



xyplot(new_data$premium ~ new_data$late_payment , group = new_data$no_default, grid = T, asp = 1)

xyplot(train_set$dim_1 ~ train_set$dim_5 , group = train_set$car, grid = T, asp = 1)

new_data$by
```

